Fitting metabolic models to dissolved oxygen data: The Estuarine BAyesian Single-station Estimation method

### 2023-10-16 manu-draft.qmd
remove priorsumcomp fig

@fig-priorsumcomp provides an alternative summary of how the prior distributions influenced the ability of EBASE to reproduce the synthetic results. Median NSE values across all models, parameters, and each optimization period for the different subsets of the prior combinations are shown. For example, the left-most bars in top left panel with the x-axis value for the $a$ parameter shows the median NSE values for all models using a 1-day optimization period where the $\mu$ value for the $a$ prior distribution was low or high. The lines indicate the inter-quartile range of NSE values for each subset. As such, each set of points denoted by the x-axis parameters in each panel include median NSE values for all models by optimization period, but subset by the low/high $\mu$ or $\sigma$ values for each parameter. Dissolved oxygen was generally modeled with high skill for all priors and it was excluded to not skew the assessment. The patterns suggest that EBASE results are more sensitive to model optimization period as shown by higher median NSE values with longer optimization times (note the different y-axis scales, also confirmed in @fig-priorplot). Similarly, median values were much more variable for the 1 day optimization period, as shown by a mix of both positive and negative values in the interquartile ranges. Lower median NSE values were observed for higher values for $\mu$, where higher median values for were observed for higher values of $\sigma$, regardless of optimization period.

@@

![Comparison of EBASE results to a synthetic time series created with known model parameters. ($a$, $R$, and $b$). Results are compared for unique combinations of prior distributions and optimization period. The means ($\mu$) and standard deviations ($\sigma$) are evaluated as low (L) or high (H) values (see text for details) and the optimization period for EBASE is 1 day, 7 days, or 30 days. The model predictions of dissolved oxygen (DO), gross primary production (P), ecosystem respiration (R), gas exchange (D), and the $a$ parameter are evaluated based on Nash-Sutcliffe Efficiency (NSE) values between the EBASE results and synthetic values. The $b$ parameter is a fixed constant and was not evaluated. Rankings of each combination as the median NSE value across all five model outputs in a row are shown on the right, with the top three as black/bold and bottom three as black/italic.](figs/priorcomp.png){#fig-priorcomp}

![Summaries of Nash-Sutcliffe Efficiency (NSE) values across all the comparisons of EBASE and the five synthetic model outputs separated by different combinations of prior values (note the different y-axis scales). Each result on the x-axis represents the median values for which the combination prior applies. For example, the bars in the upper left panel that are farthest to the left summarize the results from @fig-priorcomp where $\mu$ is high or low for the $a$ parameter using a 1 day optimization period. Bars are the medians and lines the interquartile ranges of NSE values for each result ($n$ = 8 for each bar) .](figs/priorsumcomp.png){#fig-priorsumcomp}

>>>>

@@

![Comparison of EBASE results to a synthetic time series created with known model parameters ($a$, $R$, and $b$). Results are compared for unique combinations of prior distributions and optimization period. The means ($\mu$) and standard deviations ($\sigma$) are evaluated as low (L) or high (H) values (see text for details) and the optimization period for EBASE is 1 day, 7 days, or 30 days. The model predictions of dissolved oxygen (DO), gross primary production (P), ecosystem respiration (R), air-water oxygen flux (D), and light efficiency ($a$) are evaluated based on Nash-Sutcliffe Efficiency (NSE) values between the EBASE results and synthetic values. The $b$ parameter is a fixed constant and was not evaluated. Rankings of each combination as the median NSE value across all five model outputs in a row are shown on the right, with the top three as black/bold and bottom three as black/italic.](figs/priorcomp.png){#fig-priorcomp}
